---
title: "scRNA-seq integrative analysis: log2FC difference between Ctrl and KD cells"
author: "Jeon Lee"
date: "February 3, 2021; updated on February 5 & 6 & 10 & 19 & 23 & 26, March 12"
output: html_document
---

$~$

## 0) Environment set-up and data loading
```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
library(Seurat)
library(ggplot2)
library(Matrix)
library(plotly)
library(knitr)
library(kableExtra)
library(gplots)
library(dplyr)
library(cowplot)
library(patchwork)


setwd("put_your_working_directory_here")

#======== Read a list of cell cycle markers, from Tirosh et al, 2015 ======##
cc.genes <- readLines(con = "cell_cycle_vignette_files/regev_lab_cell_cycle_genes.txt")
# We can segregate this list into markers of G2/M phase and markers of S phase
s.genes <- cc.genes[1:43]
g2m.genes <- cc.genes[44:97]
```

$~$

#### Load and process Ctrl 0hr scRNA-seq data
```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
## Load 10X experiment data
matrix_dir = "/2020_12_02_10X14_10480_0/analysis_results/ExpectCells/Ctrl_0hr/outs/"

barcode.path <- paste0(matrix_dir, "filtered_feature_bc_matrix/barcodes.tsv.gz")
features.path <- paste0(matrix_dir, "filtered_feature_bc_matrix/features.tsv.gz")
matrix.path <- paste0(matrix_dir, "filtered_feature_bc_matrix/matrix.mtx.gz")
mat <- readMM(file = matrix.path)
feature.names = read.delim(features.path, 
                           header = FALSE,
                           stringsAsFactors = FALSE)
barcode.names = read.delim(barcode.path, 
                           header = FALSE,
                           stringsAsFactors = FALSE)
colnames(mat) = barcode.names$V1
rownames(mat) = feature.names$V1 #feature.names$V2

# Add "_2" to the gene symbols which are shown twice in 'feature.names'
idx <- which(duplicated(feature.names$V2, fromLast=TRUE))
feature.names$V2[idx] <- paste(feature.names$V2[idx], "2", sep="_")
rownames(mat) = feature.names$V2

# Creat a Seurate Object
CTRL_0HR <- CreateSeuratObject(count = mat, min.cells = 0, min.features=1, project = "Ctrl_0hr")

# Calculate percent.mito values
mito.genes <- grep(pattern = "^MT-", x = rownames(x = CTRL_0HR@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(CTRL_0HR@assays[["RNA"]][mito.genes, ])/Matrix::colSums(CTRL_0HR@assays[["RNA"]])*100
CTRL_0HR <- AddMetaData(object = CTRL_0HR, metadata = percent.mito, col.name = "percent.mito") 

# Plot 'nGene', 'nUMI' and 'percent.mito' violin plots
VlnPlot(CTRL_0HR, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

## Cell fitering based on scatter plots of percent.mito and nFeature_RNA over nCount_RNA
plot1 <- FeatureScatter(object = CTRL_0HR, feature1 = "nCount_RNA", feature2 = "percent.mito")
plot2 <- FeatureScatter(object = CTRL_0HR, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1+plot2 #CombinePlots(plots = list(plot1, plot2))

## Filter out cells 
upper_thr_Count <- quantile(CTRL_0HR@meta.data$nCount_RNA, 0.90)#95) # calculate value in the 95th percentile
lower_thr_Count <- quantile(CTRL_0HR@meta.data$nCount_RNA, 0.10)#05) 
upper_thr_Feature <- quantile(CTRL_0HR@meta.data$nFeature_RNA, 0.9) #0.9)
lower_thr_Feature <- quantile(CTRL_0HR@meta.data$nFeature_RNA, 0.1) #0.1)
upper_thr_pct.mito <- 10#10
lower_thr_pct.mito <- 1#1
CTRL_0HR <- subset(x=CTRL_0HR, nCount_RNA < upper_thr_Count & nCount_RNA > lower_thr_Count & nFeature_RNA < upper_thr_Feature & nFeature_RNA > lower_thr_Feature & percent.mito < upper_thr_pct.mito & percent.mito > lower_thr_pct.mito)
```

$~$

#### Load and process KD 0hr scRNA-seq data
```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
## Load 10X experiment data
matrix_dir = "/2020_12_02_10X14_10480_0/analysis_results/ExpectCells/KD_0hr/outs/"

barcode.path <- paste0(matrix_dir, "filtered_feature_bc_matrix/barcodes.tsv.gz")
features.path <- paste0(matrix_dir, "filtered_feature_bc_matrix/features.tsv.gz")
matrix.path <- paste0(matrix_dir, "filtered_feature_bc_matrix/matrix.mtx.gz")
mat <- readMM(file = matrix.path)
feature.names = read.delim(features.path, 
                           header = FALSE,
                           stringsAsFactors = FALSE)
barcode.names = read.delim(barcode.path, 
                           header = FALSE,
                           stringsAsFactors = FALSE)
colnames(mat) = barcode.names$V1
rownames(mat) = feature.names$V1 #feature.names$V2

# Add "_2" to the gene symbols which are shown twice in 'feature.names'
idx <- which(duplicated(feature.names$V2, fromLast=TRUE))
feature.names$V2[idx] <- paste(feature.names$V2[idx], "2", sep="_")
rownames(mat) = feature.names$V2

# Creat a Seurate Object
KD_0HR <- CreateSeuratObject(count = mat, min.cells = 0, min.features=1, project = "KD_0hr")

# Calculate percent.mito values
mito.genes <- grep(pattern = "^MT-", x = rownames(x = KD_0HR@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(KD_0HR@assays[["RNA"]][mito.genes, ])/Matrix::colSums(KD_0HR@assays[["RNA"]])*100
KD_0HR <- AddMetaData(object = KD_0HR, metadata = percent.mito, col.name = "percent.mito") 

# Plot 'nGene', 'nUMI' and 'percent.mito' violin plots
VlnPlot(KD_0HR, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

## Cell fitering based on scatter plots of percent.mito and nFeature_RNA over nCount_RNA
plot1 <- FeatureScatter(object = KD_0HR, feature1 = "nCount_RNA", feature2 = "percent.mito")
plot2 <- FeatureScatter(object = KD_0HR, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1+plot2 #CombinePlots(plots = list(plot1, plot2))

## Filter out cells 
upper_thr_Count <- quantile(KD_0HR@meta.data$nCount_RNA, 0.90)#95) # calculate value in the 95th percentile
lower_thr_Count <- quantile(KD_0HR@meta.data$nCount_RNA, 0.10)#05) 
upper_thr_Feature <- quantile(KD_0HR@meta.data$nFeature_RNA, 0.9) #0.9)
lower_thr_Feature <- quantile(KD_0HR@meta.data$nFeature_RNA, 0.1) #0.1)
upper_thr_pct.mito <- 10#10
lower_thr_pct.mito <- 1#1
KD_0HR <- subset(x=KD_0HR, nCount_RNA < upper_thr_Count & nCount_RNA > lower_thr_Count & nFeature_RNA < upper_thr_Feature & nFeature_RNA > lower_thr_Feature & percent.mito < upper_thr_pct.mito & percent.mito > lower_thr_pct.mito)
```

$~$

#### Load and process Ctrl 4hr scRNA-seq data
```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
## Load 10X experiment data
matrix_dir = "/2020_12_03_10X14_10480_0/analysis_results/ExpectCells/Ctrl_4hr/outs/"

barcode.path <- paste0(matrix_dir, "filtered_feature_bc_matrix/barcodes.tsv.gz")
features.path <- paste0(matrix_dir, "filtered_feature_bc_matrix/features.tsv.gz")
matrix.path <- paste0(matrix_dir, "filtered_feature_bc_matrix/matrix.mtx.gz")
mat <- readMM(file = matrix.path)
feature.names = read.delim(features.path, 
                           header = FALSE,
                           stringsAsFactors = FALSE)
barcode.names = read.delim(barcode.path, 
                           header = FALSE,
                           stringsAsFactors = FALSE)
colnames(mat) = barcode.names$V1
rownames(mat) = feature.names$V1 #feature.names$V2

# Add "_2" to the gene symbols which are shown twice in 'feature.names'
idx <- which(duplicated(feature.names$V2, fromLast=TRUE))
feature.names$V2[idx] <- paste(feature.names$V2[idx], "2", sep="_")
rownames(mat) = feature.names$V2

# Creat a Seurate Object
CTRL_4HR <- CreateSeuratObject(count = mat, min.cells = 0, min.features=1, project = "Ctrl_4hr")

# Calculate percent.mito values
mito.genes <- grep(pattern = "^MT-", x = rownames(x = CTRL_4HR@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(CTRL_4HR@assays[["RNA"]][mito.genes, ])/Matrix::colSums(CTRL_4HR@assays[["RNA"]])*100
CTRL_4HR <- AddMetaData(object = CTRL_4HR, metadata = percent.mito, col.name = "percent.mito") 

# Plot 'nGene', 'nUMI' and 'percent.mito' violin plots
VlnPlot(CTRL_4HR, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

## Cell fitering based on scatter plots of percent.mito and nFeature_RNA over nCount_RNA
plot1 <- FeatureScatter(object = CTRL_4HR, feature1 = "nCount_RNA", feature2 = "percent.mito")
plot2 <- FeatureScatter(object = CTRL_4HR, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1+plot2 #CombinePlots(plots = list(plot1, plot2))

## Filter out cells 
upper_thr_Count <- quantile(CTRL_0HR@meta.data$nCount_RNA, 0.90)#95) # calculate value in the 95th percentile
lower_thr_Count <- quantile(CTRL_0HR@meta.data$nCount_RNA, 0.10)#05) 
upper_thr_Feature <- quantile(CTRL_0HR@meta.data$nFeature_RNA, 0.9) #0.9)
lower_thr_Feature <- quantile(CTRL_0HR@meta.data$nFeature_RNA, 0.1) #0.1)
upper_thr_pct.mito <- 10#10
lower_thr_pct.mito <- 1#1
CTRL_4HR <- subset(x=CTRL_4HR, nCount_RNA < upper_thr_Count & nCount_RNA > lower_thr_Count & nFeature_RNA < upper_thr_Feature & nFeature_RNA > lower_thr_Feature & percent.mito < upper_thr_pct.mito & percent.mito > lower_thr_pct.mito)
```

$~$

#### Load and process KD 4hr scRNA-seq data
```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
## Load 10X experiment data
matrix_dir = "/2020_12_03_10X14_10480_0/analysis_results/ExpectCells/KD_4hr/outs/"

barcode.path <- paste0(matrix_dir, "filtered_feature_bc_matrix/barcodes.tsv.gz")
features.path <- paste0(matrix_dir, "filtered_feature_bc_matrix/features.tsv.gz")
matrix.path <- paste0(matrix_dir, "filtered_feature_bc_matrix/matrix.mtx.gz")
mat <- readMM(file = matrix.path)
feature.names = read.delim(features.path, 
                           header = FALSE,
                           stringsAsFactors = FALSE)
barcode.names = read.delim(barcode.path, 
                           header = FALSE,
                           stringsAsFactors = FALSE)
colnames(mat) = barcode.names$V1
rownames(mat) = feature.names$V1 #feature.names$V2

# Add "_2" to the gene symbols which are shown twice in 'feature.names'
idx <- which(duplicated(feature.names$V2, fromLast=TRUE))
feature.names$V2[idx] <- paste(feature.names$V2[idx], "2", sep="_")
rownames(mat) = feature.names$V2

# Creat a Seurate Object
KD_4HR <- CreateSeuratObject(count = mat, min.cells = 0, min.features=1, project = "KD_4hr")

# Calculate percent.mito values
mito.genes <- grep(pattern = "^MT-", x = rownames(x = KD_4HR@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(KD_4HR@assays[["RNA"]][mito.genes, ])/Matrix::colSums(KD_4HR@assays[["RNA"]])*100
KD_4HR <- AddMetaData(object = KD_4HR, metadata = percent.mito, col.name = "percent.mito") 

# Plot 'nGene', 'nUMI' and 'percent.mito' violin plots
VlnPlot(KD_4HR, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

## Cell fitering based on scatter plots of percent.mito and nFeature_RNA over nCount_RNA
plot1 <- FeatureScatter(object = KD_4HR, feature1 = "nCount_RNA", feature2 = "percent.mito")
plot2 <- FeatureScatter(object = KD_4HR, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1+plot2 #CombinePlots(plots = list(plot1, plot2))

## Filter out cells 
upper_thr_Count <- quantile(KD_4HR@meta.data$nCount_RNA, 0.90)#95) # calculate value in the 95th percentile
lower_thr_Count <- quantile(KD_4HR@meta.data$nCount_RNA, 0.10)#05) 
upper_thr_Feature <- quantile(KD_4HR@meta.data$nFeature_RNA, 0.9) #0.9)
lower_thr_Feature <- quantile(KD_4HR@meta.data$nFeature_RNA, 0.1) #0.1)
upper_thr_pct.mito <- 10#10
lower_thr_pct.mito <- 1#1
KD_4HR <- subset(x=KD_4HR, nCount_RNA < upper_thr_Count & nCount_RNA > lower_thr_Count & nFeature_RNA < upper_thr_Feature & nFeature_RNA > lower_thr_Feature & percent.mito < upper_thr_pct.mito & percent.mito > lower_thr_pct.mito)
```

$~$

## 1) Integration of Ctrl scRNA-seq data (Ctrl_0hr and Ctrl_4hr)
```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
options(future.globals.maxSize = 8000 * 1024^2) # set to 8GB

# Generate a list variable of Seurat objects
Ctrl.list=list()
Ctrl.list[[1]] <- CTRL_0HR
Ctrl.list[[2]] <- CTRL_4HR
names(Ctrl.list) <- c("CTRL_0HR", "CTRL_4HR")

for (i in 1:length(Ctrl.list)) {
    Ctrl.list[[i]] <- NormalizeData(Ctrl.list[[i]], verbose = TRUE)
    Ctrl.list[[i]] <- CellCycleScoring(Ctrl.list[[i]], g2m.features=g2m.genes, s.features=s.genes)
    Ctrl.list[[i]] <- SCTransform(Ctrl.list[[i]], vars.to.regress = c("percent.mito"))
    ##Ctrl.list[[i]] <- SCTransform(Ctrl.list[[i]], vars.to.regress = c("S.Score","G2M.Score"))
}

Ctrl.features <- SelectIntegrationFeatures(object.list = Ctrl.list, nfeatures = 7500)#5000) #3000)
Ctrl.list <- PrepSCTIntegration(object.list = Ctrl.list, anchor.features = Ctrl.features, verbose = FALSE)

Ctrl.anchors <- FindIntegrationAnchors(object.list = Ctrl.list, normalization.method = "SCT", anchor.features = Ctrl.features, verbose = FALSE)
Ctrl.integrated <- IntegrateData(anchorset = Ctrl.anchors, normalization.method = "SCT", verbose = FALSE)

head(x = Ctrl.integrated@meta.data)

rm(list=c("Ctrl.list","Ctrl.anchors"))
```

$~$

#### Clustering analysis for the integrated Ctrl scRNA-seq data
```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
# specify that we will perform downstream analysis on the corrected data note that the original
# unmodified data still resides in the 'RNA' assay
DefaultAssay(Ctrl.integrated) <- "integrated"

# Run the standard workflow for visualization and clustering
Ctrl.integrated <- RunPCA(Ctrl.integrated, npcs = 30, verbose = FALSE)
Ctrl.integrated <- RunUMAP(Ctrl.integrated, reduction = "pca", dims = 1:30)

set.seed(12345)
Ctrl.integrated <- FindNeighbors(Ctrl.integrated, reduction = "pca", dims = 1:30)
Ctrl.integrated <- FindClusters(Ctrl.integrated, resolution = 0.75)

# Visualization
p1 <- DimPlot(Ctrl.integrated, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(Ctrl.integrated, reduction = "umap", label = TRUE, repel = TRUE)
p1+p2
DimPlot(Ctrl.integrated, reduction = "umap", split.by = "orig.ident")

## CD69 expression levels
DefaultAssay(Ctrl.integrated) <- "RNA"
FeaturePlot(Ctrl.integrated, features = c("CD69"), split.by = "orig.ident", max.cutoff = 3, cols = c("grey", "red"))
```

$~$

#### (optional) Trajectory analysis
```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=5}
library(SeuratWrappers)
library(monocle3)

Idents(Ctrl.integrated) <- "orig.ident"
Ctrl.integrated.sub <- subset(Ctrl.integrated, idents = c("Ctrl_4hr"))

Ctrl.cds <- as.cell_data_set(Ctrl.integrated.sub)
Ctrl.cds <- cluster_cells(cds = Ctrl.cds, reduction_method = "UMAP")
Ctrl.cds <- learn_graph(Ctrl.cds, use_partition = FALSE, close_loop = FALSE)

# cell with the highest ADAP1
idx <- which.max(assay(Ctrl.cds)[rownames(assay(Ctrl.cds))=="CD69",])

# order cells
Ctrl.cds <- order_cells(Ctrl.cds, reduction_method = "UMAP", root_cells = names(idx)) #NULL

# plot trajectories colored by pseudotime
plot_cells(
  cds = Ctrl.cds,
  color_cells_by = "pseudotime",
  show_trajectory_graph = TRUE
)
```

$~$

#### Marker discovery for Ctrl_4hr vs. Ctrl_0hr cells
```{r message=FALSE, warning=FALSE, fig.width=8, fig.height=12}
#DefaultAssay(Ctrl.integrated) <- "RNA"
#Ctrl.integrated <- NormalizeData(Ctrl.integrated, verbose = FALSE)

Ctrl.integrated <- SetIdent(object = Ctrl.integrated, value = "orig.ident")
Ctrl.integrated.markers <- FindMarkers(Ctrl.integrated, ident.1 = "Ctrl_4hr", assay="RNA", only.pos = FALSE, min.pct = 0.1, logfc.threshold = 0.25) #0.10)#25) 
#Ctrl.integrated <- SetIdent(object = Ctrl.integrated, value = "seurat_clusters")
#Ctrl.integrated.markers <- FindMarkers(Ctrl.integrated, ident.1 = c(3), only.pos = FALSE, min.pct = 0.1, logfc.threshold = 0.5) 


Ctrl.UP.sig_markers <- Ctrl.integrated.markers %>% filter(p_val_adj < 0.05) %>% top_n(n = 50, wt = avg_log2FC)
Ctrl.DN.sig_markers <- Ctrl.integrated.markers %>% filter(p_val_adj < 0.05) %>% top_n(n = -50, wt = avg_log2FC)
Ctrl.sig_markers <- bind_rows(Ctrl.UP.sig_markers, Ctrl.DN.sig_markers)

#DefaultAssay(Ctrl.integrated) <- "integrated"
#DoHeatmap(Ctrl.integrated, features = rownames(Ctrl.sig_markers), size=2) + NoLegend() + theme(text = element_text(size = 8))
```
$~$

```{r message=FALSE, warning=FALSE, fig.width=12, fig.height=50}
#Ctrl.UP.sig_markers
Ctrl.up_sig_genes <- rownames(Ctrl.UP.sig_markers)

Ctrl.up_plots <- VlnPlot(Ctrl.integrated, features = Ctrl.up_sig_genes, assay="RNA", group.by = "orig.ident", pt.size = 0.5, combine = FALSE)
wrap_plots(plots=Ctrl.up_plots, ncol=3)
```

```{r message=FALSE, warning=FALSE, fig.width=12, fig.height=50}
#Ctrl.DN.sig_markers
Ctrl.down_sig_genes <- rownames(Ctrl.DN.sig_markers)

Ctrl.down_plots <- VlnPlot(Ctrl.integrated, features = Ctrl.down_sig_genes, assay="RNA", group.by = "orig.ident", pt.size = 0.5, combine = FALSE)
wrap_plots(plots=Ctrl.down_plots, ncol=3)
```


$~$

## 2) Integration of KD scRNA-seq data (KD_0hr and _4hr)
```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
options(future.globals.maxSize = 8000 * 1024^2) # set to 8GB

# Generate a list variable of Seurat objects
KD.list=list()
KD.list[[1]] <- KD_0HR
KD.list[[2]] <- KD_4HR
names(KD.list) <- c("KD_0HR", "KD_4HR")

for (i in 1:length(KD.list)) {
    KD.list[[i]] <- NormalizeData(KD.list[[i]], verbose = TRUE)
    KD.list[[i]] <- CellCycleScoring(KD.list[[i]], g2m.features=g2m.genes, s.features=s.genes)
    KD.list[[i]] <- SCTransform(KD.list[[i]], vars.to.regress = c("percent.mito"))
    #KD.list[[i]] <- SCTransform(KD.list[[i]], vars.to.regress = c("S.Score","G2M.Score"))
}

KD.features <- SelectIntegrationFeatures(object.list = KD.list, nfeatures = 7500) #5000) #3000)
KD.list <- PrepSCTIntegration(object.list = KD.list, anchor.features = KD.features, verbose = FALSE)

KD.anchors <- FindIntegrationAnchors(object.list = KD.list, normalization.method = "SCT", anchor.features = KD.features, verbose = FALSE)
KD.integrated <- IntegrateData(anchorset = KD.anchors, normalization.method = "SCT", verbose = FALSE)

head(x = KD.integrated@meta.data)

rm(list=c("KD.list","KD.anchors"))
```

$~$

#### Clustering analysis for the integrated KD scRNA-seq data
```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
# specify that we will perform downstream analysis on the corrected data note that the original
# unmodified data still resides in the 'RNA' assay
DefaultAssay(KD.integrated) <- "integrated"

# Run the standard workflow for visualization and clustering
KD.integrated <- RunPCA(KD.integrated, npcs = 50, verbose = FALSE)
KD.integrated <- RunUMAP(KD.integrated, reduction = "pca", dims = 1:50)

set.seed(12345)
KD.integrated <- FindNeighbors(KD.integrated, reduction = "pca", dims = 1:50)
KD.integrated <- FindClusters(KD.integrated, resolution = 0.75)

# Visualization
p1 <- DimPlot(KD.integrated, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(KD.integrated, reduction = "umap", label = TRUE, repel = TRUE)
p1+p2
DimPlot(KD.integrated, reduction = "umap", split.by = "orig.ident")

## CD69 expression levels
DefaultAssay(KD.integrated) <- "RNA"
FeaturePlot(KD.integrated, features = c("CD69"), split.by = "orig.ident", max.cutoff = 3, cols = c("grey", "red"))
```

$~$

#### (optional) Trajectory analysis
```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=5}
library(SeuratWrappers)
library(monocle3)

Idents(KD.integrated) <- "orig.ident"
KD.integrated.sub <- subset(KD.integrated, idents = c("KD_4hr"))

KD.cds <- as.cell_data_set(KD.integrated.sub)
KD.cds <- cluster_cells(cds = KD.cds, reduction_method = "UMAP")
KD.cds <- learn_graph(KD.cds, use_partition = FALSE, close_loop = FALSE)

# cell with the highest ADAP1
idx <- which.max(assay(KD.cds)[rownames(assay(KD.cds))=="CD69",])

# order cells
KD.cds <- order_cells(KD.cds, reduction_method = "UMAP", root_cells = names(idx)) #NULL

# plot trajectories colored by pseudotime
plot_cells(
  cds = KD.cds,
  color_cells_by = "pseudotime",
  show_trajectory_graph = TRUE
)
```

$~$

#### Marker discovery for KD_4hr vs. KD_0hr cells
```{r message=FALSE, warning=FALSE, fig.width=8, fig.height=12}
#DefaultAssay(KD.integrated) <- "RNA"
#KD.integrated <- NormalizeData(KD.integrated, verbose = FALSE)

KD.integrated <- SetIdent(object = KD.integrated, value = "orig.ident")
KD.integrated.markers <- FindMarkers(KD.integrated, ident.1 = "KD_4hr", assay="RNA", only.pos = FALSE, min.pct = 0.1, logfc.threshold = 0.25)#0.10)#25)
#KD.integrated <- SetIdent(object = KD.integrated, value = "seurat_clusters")
#KD.integrated.markers <- FindMarkers(KD.integrated, ident.1 = c(3), only.pos = FALSE, min.pct = 0.1, logfc.threshold = 0.5) 

KD.UP.sig_markers <- KD.integrated.markers %>% filter(p_val_adj < 0.05) %>% top_n(n = 50, wt = avg_log2FC)
KD.DN.sig_markers <- KD.integrated.markers %>% filter(p_val_adj < 0.05) %>% top_n(n = -50, wt = avg_log2FC)
KD.sig_markers <- bind_rows(KD.UP.sig_markers, KD.DN.sig_markers)

#DefaultAssay(KD.integrated) <- "integrated"
#DoHeatmap(KD.integrated, features = rownames(KD.sig_markers), size=2) + NoLegend() + theme(text = element_text(size = 8))
```
$~$

```{r message=FALSE, warning=FALSE, fig.width=12, fig.height=50}
#KD.UP.sig_markers
KD.up_sig_genes <- rownames(KD.UP.sig_markers)

KD.up_plots <- VlnPlot(KD.integrated, features = KD.up_sig_genes, assay="RNA", group.by = "orig.ident", pt.size = 0.5, combine = FALSE)
wrap_plots(plots=KD.up_plots, ncol=3)
```

```{r message=FALSE, warning=FALSE, fig.width=12, fig.height=50}
#KD.DN.sig_markers
KD.down_sig_genes <- rownames(KD.DN.sig_markers)

KD.down_plots <- VlnPlot(KD.integrated, features = KD.down_sig_genes, assay="RNA", group.by = "orig.ident", pt.size = 0.5, combine = FALSE)
wrap_plots(plots=KD.down_plots, ncol=3)
```

$~$


## 3) Log2FC scatter plot: Ctrl vs. KD
```{r message=FALSE, warning=FALSE, fig.width=8, fig.height=8}
markers.merged <- merge(Ctrl.integrated.markers, KD.integrated.markers, by=0, all=TRUE)
colnames(markers.merged) <- gsub(".x", ".CTRL", colnames(markers.merged))
colnames(markers.merged) <- gsub(".y", ".KD", colnames(markers.merged))

plot_ly(data=markers.merged, x = ~avg_log2FC.CTRL, y = ~avg_log2FC.KD, type="scatter", text = markers.merged$Row.names) %>% add_lines(x =  ~avg_log2FC.CTRL, y = ~avg_log2FC.CTRL) %>% layout(showlegend = FALSE)

```

$~$

```{r message=FALSE, warning=FALSE, fig.width=12, fig.height=10}
#genes.of.interest
genes.of.interest <- c("CD69", "MYC", "GZMB", "LTA", "CD40LG", "RGCC", "NR4A1", "TNF", "IL2", "CSF2")

interest_plots_ctrl <- VlnPlot(Ctrl.integrated, features = genes.of.interest, assay="RNA", group.by = "orig.ident", pt.size = 0.0, combine = FALSE)
wrap_plots(plots=interest_plots_ctrl, ncol=3)

interest_plots_kd <- VlnPlot(KD.integrated, features = genes.of.interest, assay="RNA", group.by = "orig.ident", pt.size = 0.0, combine = FALSE)
wrap_plots(plots=interest_plots_kd, ncol=3)
```

$~$

```{r message=FALSE, warning=FALSE, fig.width=8, fig.height=3.5}
Ctrl.integrated <- AddModuleScore(object = Ctrl.integrated, features = genes.of.interest, name = "cell_active_score")
FeaturePlot(object = Ctrl.integrated, features = "cell_active_score1", split.by = "orig.ident", min.cutoff=0.1, max.cutoff = 3.5, cols = c("grey", "red"))

KD.integrated <- AddModuleScore(object = KD.integrated, features = genes.of.interest, name = "cell_active_score")
FeaturePlot(object = KD.integrated, features = "cell_active_score1", split.by = "orig.ident", min.cutoff=0.1, max.cutoff = 3.5, cols = c("grey", "red"))
```

$~$

### ==End of Report==
